### 2016-12-09 Docs/Communication/discrepancy.tex
skeleton_slides

\documentclass[english,11pt]{beamer}

\DeclareMathOperator{\Cov}{Cov}
\DeclareMathOperator{\Var}{Var}
\DeclareMathOperator{\E}{\mathbb{E}}
\DeclareMathOperator{\Proba}{\mathbb{P}}

\newcommand{\Covb}[2]{\ensuremath{\Cov\!\left[#1,#2\right]}}
\newcommand{\Eb}[1]{\ensuremath{\E\!\left[#1\right]}}
\newcommand{\Pb}[1]{\ensuremath{\Proba\!\left[#1\right]}}
\newcommand{\Varb}[1]{\ensuremath{\Var\!\left[#1\right]}}

% norm
\newcommand{\norm}[1]{\| #1 \|}

\newcommand{\indep}{\rotatebox[origin=c]{90}{$\models$}}

\usepackage{mathptmx,amsmath,amssymb,graphicx,bibentry,bbm,babel,ragged2e}

\makeatletter

\newcommand{\noun}[1]{\textsc{#1}}
\newcommand{\jitem}[1]{\item \begin{justify} #1 \end{justify} \vfill{}}
\newcommand{\sframe}[2]{\frame{\frametitle{#1} #2}}

\newenvironment{centercolumns}{\begin{columns}[c]}{\end{columns}}
%\newenvironment{jitem}{\begin{justify}\begin{itemize}}{\end{itemize}\end{justify}}

\usetheme{Warsaw}
\setbeamertemplate{footline}[text line]{}
\setbeamercolor{structure}{fg=purple!50!blue, bg=purple!50!blue}

\setbeamersize{text margin left=15pt,text margin right=15pt}

\setbeamercovered{transparent}

\@ifundefined{showcaptionsetup}{}{%
 \PassOptionsToPackage{caption=false}{subfig}}
\usepackage{subfig}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

\makeatother

\begin{document}

\title{A Discrepancy-based Framework to Compare Robustness between Multi-Attribute Evaluations}

\author{J.~Raimbault$^{1,2}$\\
\texttt{juste.raimbault@parisgeo.cnrs.fr}
}

\institute{$^{1}$UMR CNRS 8504 G{\'e}ographie-cit{\'e}s\\
$^{2}$UMR-T IFSTTAR 9403 LVMT\\
}

\date{CSDM 2016 - Paris\\\smallskip
14th December 2016
}

\frame{\maketitle}

%%%%%%%%%%%%%%%%%
\section{Introduction}
%%%%%%%%%%%%%%%%%

\sframe{Complex systems}{

% introduction with pictures/anecdote

}

\sframe{Multi-attribute Evaluation in Complex Socio-technical Systems}{

\justify

}

\sframe{Robustness of Evaluations}{

\justify

}

\sframe{Towards a Framework}{

\justify

\textbf{Research Objective : } \textit{}

% put rationales here

\bigskip

$\rightarrow$ 
\medskip

$\rightarrow$

}

%%%%%%%%%%%%%%%%%
\section{Framework Description}
%%%%%%%%%%%%%%%%%

\sframe{Intuitive Description}{

}

\sframe{Formal Description (I)}{

}

\sframe{Formal Description (II)}{

}

%%%%%%%%%%%%%%%%%
\section{Results}
%%%%%%%%%%%%%%%%%

\sframe{Implementation on Synthetic Data}{

% summary, data description, main result (summary of table)

}

\sframe{Metropolitan Segregation}{

% indicators, data

}

\sframe{Metropolitan Segregation}{

% segreg maps

}

\sframe{Metropolitan Segregation}{

% missing data

}

%%%%%%%%%%%%%%%%%
\section{Discussion}
%%%%%%%%%%%%%%%%%

\sframe{Applicability}{

\justify

}

\sframe{Possible Developments}{

\justify

}

\sframe{Conclusion}{

$\rightarrow$

\medskip

\bigskip
\bigskip
\bigskip

\footnotesize{ - All code available at \texttt{https://github.com/JusteRaimbault/RobustnessDiscrepancy}

\medskip

 - Paper preprint available at \texttt{http://arxiv.org/abs/}
}

}

%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[allowframebreaks]
\frametitle{References}
\bibliographystyle{apalike}
\bibliography{/Users/Juste/Documents/ComplexSystems/CityNetwork/Biblio/Bibtex/CityNetwork,biblio}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}